---
title: "meta-data"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r include, echo=F, warning=F}

suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggvenn))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(MLmetrics))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(tidyr))

source("../src/novel-discovery/obtain-data.R")
source("../src/novel-discovery/load-data.R")

```

```{r, echo=F}

prop.shower <- function(prop.data, threshold, show.t=T) {
  prop.sel <- prop.data[,2:ncol(prop.data)] %>% rowMeans
  prop.sel <- data.frame("feature"=prop.data[,1],
                         "prop"=prop.sel)
  p <- ggplot(prop.sel, aes(x=feature, y=prop)) +
    coord_flip() +
    geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=45, hjust=1, size=10),
          axis.text.y=element_text(size=6)) +
    ylab("proportion") +
    xlab("")
  if (show.t) {
    p <- p + geom_hline(yintercept=threshold, colour="red", linetype=2)
  }
  print(p)
  
  sel.list <- prop.sel$feature[prop.sel$prop >= threshold]
  if (show.t) {
    cat("\n\n Selected features are: ", paste(sel.list, collapse=", "))
  }
  sel.list
}

```

```{r, echo=F}

obtain.metrics <- function(df) {
  # df contains true and pred as columns.
  pred.class <- ifelse(df$pred >= 0.5, "Positive", "Negative")
  auc.val <- AUC(df$pred, ifelse(df$true == "Positive", 1, 0))
  sens.val <- Sensitivity(df$true, pred.class, positive="Positive")
  spec.val <- Specificity(df$true, pred.class, positive="Positive")
  acc.val <- Accuracy(df$true, pred.class)
  list("Accuracy"=acc.val,
       "Sensitivity"=sens.val,
       "Specificity"=spec.val,
       "AUC"=auc.val)
}

obtain.bootstrap.metrics <- function(boot.runs) {
  metric.df <- data.frame(matrix(data=NA, nrow=0, ncol=4))
  pred.df <- data.frame(matrix(data=NA, nrow=0, ncol=2))
  colnames(pred.df) <- c("pred", "true")
  for (i in 1:length(boot.runs)) {
    # fold.1.i <- which(boot.runs[[i]]$fold == 2)
    # metric.boot <- obtain.metrics(boot.runs[[i]][fold.1.i,])
    metric.boot <- obtain.metrics(boot.runs[[i]])
    metric.df <- rbind(metric.df, metric.boot)
    pred.df <- rbind(pred.df, 
                     data.frame("pred"=boot.runs[[i]]$pred,
                                "true"=boot.runs[[i]]$true))
    # pred.df <- rbind(pred.df, 
    #                  data.frame("pred"=boot.runs[[i]]$pred[fold.1.i],
    #                             "true"=boot.runs[[i]]$true[fold.1.i]))
  }
  metric.mean.df <- round(metric.df %>% colMeans, 3)
  metric.quantiles <- metric.df %>% apply(2, \(x) {
    metric.q <- quantile(x, c(0.025, 0.975))
    paste0("(", round(metric.q[1], 3), "-", round(metric.q[2], 3), ")")
  })
  
  metric.mean.df <- rbind(metric.mean.df, metric.quantiles)
  
  print(metric.mean.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered"))
  
  roc.curve <- prediction(pred.df$pred, pred.df$true, label.ordering=c("Negative", "Positive"))
  pred.roc <- performance(roc.curve, "tpr", "fpr")
  perf.df <- data.frame(
    "fpr"=unlist(pred.roc@x.values),
    "tpr"=unlist(pred.roc@y.values)
  )
  p <- ggplot(perf.df, aes(fpr, tpr)) +
    geom_line() +
    geom_abline(slope=1, intercept=0, linetype=2, colour="black")
  print(p)
  
  metric.mean.df[1,,drop=F]
}

```

```{r, echo=F}

p.df <- data.frame(matrix(data=NA, nrow=0, ncol=2))
colnames(p.df) <- c("Name", "p.value")

```

```{r, echo=F}

data <- load.data(path.to.root="..")
```

```{r, echo=F}

val.data.6 <- read.xlsx("../data/Including Day4_Validation chorts_260324.xlsx", sheet=8)
val.data.6 <- val.data.6$Seq_Sample_ID
train.6.i <- which(!data$meta.data$seq.id %in% val.data.6)

train.6 <- list("meta.data"=data$meta.data[train.6.i,],
                "x"=data$x[train.6.i,])
# Add stage to data.
train.6$meta.data$stage <- ifelse(grepl("AA", train.6$meta.data$Emb_stage), 1, 
                                 ifelse(grepl("AB|BA", train.6$meta.data$Emb_stage), 2,
                                        ifelse(grepl("BB", train.6$meta.data$Emb_stage), 3, 4)))
train.6$meta.data$stage <- factor(train.6$meta.data$stage, levels=c(1, 2, 3, 4))

all.train <- list("meta.data"=data$meta.data,
                  "x"=data$x)

```

```{r, echo=F}

get.reg.data <- function(meta.data, col.interest) {
  morph.train <- data.frame("hCG.result"=meta.data$hCG_Result)
  morph.train[col.interest] <- meta.data[col.interest]

  min.mod <- glm(hCG.result~1, morph.train, family=binomial)
  max.mod <- glm(hCG.result~., morph.train, family=binomial)
  print(summary(max.mod))
  
  print(anova(min.mod, max.mod, test="LRT"))
  list("Name"=col.interest, "p.value"=anova(min.mod, max.mod, test="LRT")[2, "Pr(>Chi)"])
}

```

# Introduction

Here, I conduct statistical analysis on different meta-data. I have summaries of the p-values in a table under each section. It is called
"pvalue summary". There are also printouts of the statistical analysis
if you want to see effect sizes and other statistical information, like
deviance of the model, etc.

In summary, it seems that G3 and Age are the only features that are 
significantly different in the negative hcg group vs positive hcg group.

# Meta data analysis using the entire dataset

Here, I have used the entire 75 patients for statistical analysis.

## Age

```{r, echo=F}
p.df <- rbind(p.df, get.reg.data(all.train$meta.data, "Age"))

```

## Morphological data

### G1

```{r, echo=F}
p.df <- rbind(p.df, get.reg.data(all.train$meta.data, "G1"))

```

G1 doesn't seem significant.

### G2

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(all.train$meta.data, "G2"))

```

G2 doesn't seem significant.

### G3

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(all.train$meta.data, "G3"))

```

## IVF_center

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "IVF_center"))

```

The IVF center is likely not important. However, I will note that we only have ~10
samples from the Uppsala site - the variance will be very large here, likely
contributing to much higher p-values too.


## Attempt_nr

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "Attempt_nr"))

```

## Culture_Day

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "Culture_Day"))

```

## Hormonal_Treatment

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "Hormonal_Treatment"))

```

## Method

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "Method"))

```

## Media

```{r, echo=F}

p.df <- rbind(p.df, get.reg.data(data$meta.data, "Media"))

```

## pvalue summary

```{r, echo=F, results="asis"}

p.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered")

```

# Meta data analysis using the training set of val-6

Here, I've conducted statistical analysis on the 65 patients in the
training set of val-6.

```{r, echo=F}

val.6.p.df <- data.frame(matrix(data=NA, nrow=0, ncol=2))
colnames(val.6.p.df) <- c("Name", "p.value")

```

## Age

```{r, echo=F}
val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "Age"))

```


## Morphological data

### G1

```{r, echo=F}
val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "G1"))

```

G1 doesn't seem significant.

### G2

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "G2"))

```

G2 doesn't seem significant.

### G3

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "G3"))

```

By the regression, it appears that G3 is impactful to the
performance. I'll include this in our regression
modelling.


## IVF_center

Please note that val6 is separated by IVF_center. I.e. the column
associated with the ivf centre is not changing in the training set.
Hence, I have not done the analysis here.

## Attempt_nr

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "Attempt_nr"))

```

## Culture_Day

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "Culture_Day"))

```

## Hormonal_Treatment

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "Hormonal_Treatment"))

```

## Method

```{r, echo=F}

val.6.p.df <- rbind(val.6.p.df, get.reg.data(train.6$meta.data, "Method"))

```

## pvalue summary

```{r, echo=F, results="asis"}

val.6.p.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered")

```
