---
title: "val-6-models"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r include, echo=F, warning=F}
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(MLmetrics))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(org.Hs.eg.db))

source("../src/novel-discovery/obtain-data.R")
source("../src/novel-discovery/load-data.R")

```

```{r, echo=F}

prop.shower <- function(prop.data, threshold, show.t=T) {
  prop.sel <- prop.data[,2:ncol(prop.data)] %>% rowMeans
  prop.sel <- data.frame("feature"=prop.data[,1],
                         "prop"=prop.sel)
  p <- ggplot(prop.sel, aes(x=feature, y=prop)) +
    coord_flip() +
    geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=45, hjust=1, size=10),
          axis.text.y=element_text(size=6)) +
    ylab("proportion") +
    xlab("")
  if (show.t) {
    p <- p + geom_hline(yintercept=threshold, colour="red", linetype=2)
  }
  print(p)
  
  sel.list <- prop.sel$feature[prop.sel$prop >= threshold]
  if (show.t) {
    cat("\n\n Selected features are: ", paste(sel.list, collapse=", "))
  }
  sel.list
}

```

```{r, echo=F}

obtain.metrics <- function(df, label.ordering) {
  # df contains true and pred as columns.
  pred.class <- ifelse(df$pred >= 0.5, label.ordering[2], label.ordering[1])
  auc.val <- AUC(df$pred, ifelse(df$true == label.ordering[2], 1, 0))
  sens.val <- Sensitivity(df$true, pred.class, positive=label.ordering[2])
  spec.val <- Specificity(df$true, pred.class, positive=label.ordering[2])
  acc.val <- Accuracy(df$true, pred.class)
  list("Accuracy"=acc.val,
       "Sensitivity"=sens.val,
       "Specificity"=spec.val,
       "AUC"=auc.val)
}

obtain.bootstrap.metrics <- function(boot.runs, label.ordering=c("Negative", "Positive")) {
  metric.df <- data.frame(matrix(data=NA, nrow=0, ncol=4))
  pred.df <- data.frame(matrix(data=NA, nrow=0, ncol=2))
  colnames(pred.df) <- c("pred", "true")
  for (i in 1:length(boot.runs)) {
    # fold.1.i <- which(boot.runs[[i]]$fold == 2)
    metric.boot <- obtain.metrics(boot.runs[[i]]$preds, label.ordering)
    metric.df <- rbind(metric.df, metric.boot)
    pred.df <- rbind(pred.df, 
                     data.frame("pred"=boot.runs[[i]]$preds$pred,
                                "true"=boot.runs[[i]]$preds$true))
  }
  metric.mean.df <- round(metric.df %>% colMeans, 3)
  metric.quantiles <- metric.df %>% apply(2, \(x) {
    metric.q <- quantile(x, c(0.025, 0.975))
    paste0("(", round(metric.q[1], 3), "-", round(metric.q[2], 3), ")")
  })
  
  metric.mean.df <- rbind(metric.mean.df, metric.quantiles)
  
  print(metric.mean.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered"))
  
  conf.factor <- factor(ifelse(pred.df$pred >= 0.5, label.ordering[2], label.ordering[1]), levels=label.ordering)
  c.mat <- confusionMatrix(conf.factor, pred.df$true)
  colnames(c.mat$table) <- paste0(colnames(c.mat$table), "-reference")
  rownames(c.mat$table) <- paste0(rownames(c.mat$table), "-prediction")
  print(c.mat$table %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered"))
  # cat("\n\n Columns are the reference, rows are the predictions \n\n")
  
  
  roc.curve <- prediction(pred.df$pred, pred.df$true, label.ordering=label.ordering)
  pred.roc <- performance(roc.curve, "tpr", "fpr")
  perf.df <- data.frame(
    "fpr"=unlist(pred.roc@x.values),
    "tpr"=unlist(pred.roc@y.values)
  )
  p <- ggplot(perf.df, aes(fpr, tpr)) +
    geom_line() +
    geom_abline(slope=1, intercept=0, linetype=2, colour="black")
  print(p)
  
  metric.mean.df[1,,drop=F]
}

```

```{r, echo=F}

sum.df <- data.frame(matrix(data=NA, nrow=0, ncol=5))
colnames(sum.df) <- c("Name", "Accuracy", "Sensitivity", "Specificity", "AUC")
internal.df <- data.frame(matrix(data=NA, nrow=0, ncol=5))
colnames(internal.df) <- c("Name", "Accuracy", "Sensitivity", "Specificity", "AUC")

```

```{r, echo=F}

data <- load.data(path.to.root="..")

```

# Introduction

Model metrics for validation set 6. Note that these metrics include
internal and external validation. I have run internal and external
validation over 10 iterations of 10 fold cross-validation. 
E.g. In my first training instance, I have trained a model on 
folds 2-10, then tested on fold 1 (internal
validation) and on validation set 6 (external validation). I then
repeated for the rest of the folds and remaining 9 iterations. Thus, I
should have 10 predictions for each sample in the training set. The 
external validation should have more - 100 predictions for each sample
in the ext-val set.

Please note that the confusion matrices
are of 10 iterations of 10 fold cross validation. I.e. for internal
10-fold cross validation, we should have 10*65=650 predictions - 10 
predictions for each sample in the training set since we have 10
iterations. For external validation,
we should have 10*10*10=1000 predictions - 10 iterations and 10 folds
for 10 patients in the validation set. The numbers of patients in each 
confusion matrix should sum to their respective values that are listed.


```{r, echo=F, results="asis"}
for (i in c("ridge", "rf")) {
  for (j in c("hCG", "lbo")) {
    for (k in c(0.05, 0.1)) {
      for (l in c("", "Age", "G1&G2&G3", "G1&G2&G3&Age", "G3", "G3&Age")) {
        
        meta.name.list <- c()
        if (grepl("G1", l)) {
          meta.name.list <- c(meta.name.list, "g1")
        }
        if (grepl("G2", l)) {
          meta.name.list <- c(meta.name.list, "g2")
        }
        if (grepl("G3", l)) {
          meta.name.list <- c(meta.name.list, "g3")
        }
        if (grepl("Age", l)) {
          meta.name.list <- c(meta.name.list, "age")
        }
        meta.name <- paste0(meta.name.list, collapse="+")

        cat("\n\n# model=",i, "filter=",k, "outcome=", j, "clin feat=", ifelse(meta.name!="", meta.name, "none"), "\n\n")
        cat("\n\n## Internal validation \n\n")
        internal.data <- readRDS(paste0("../results/kfold-panel-by-validation-split-internal/6/model-", i, "_panel-val-6-lasso_max-iter-10_k-10_meta-", l, "_filter-", k, "_class-", j, "/combine-42.rds"))
        if (j == "hCG") {
          internal.df <- rbind(internal.df, data.frame("Name"=paste0("model-", i, "_filt-", k, "_class-", j, "_clin.feat-", meta.name), obtain.bootstrap.metrics(internal.data)))
        } else if (j == "lbo") {
          internal.df <- rbind(internal.df, data.frame("Name"=paste0("model-", i, "_filt-", k, "_class-", j, "_clin.feat-", meta.name), obtain.bootstrap.metrics(internal.data, c("NPO", "Live_birth"))))
        }
        
        cat("\n\n## External validation \n\n")
        metric.data <- readRDS(paste0("../results/kfold-panel-by-validation-split/6/model-", i, "_panel-val-6-lasso_max-iter-10_k-10_meta-", l, "_filter-", k, "_class-", j, "/combine-42.rds"))
        if (j == "hCG") {
          sum.df <- rbind(sum.df, data.frame("Name"=paste0("model-", i, "_filt-", k, "_class-", j, "_clin.feat-", meta.name), obtain.bootstrap.metrics(metric.data)))
        } else if (j == "lbo") {
          sum.df <- rbind(sum.df, data.frame("Name"=paste0("model-", i, "_filt-", k, "_class-", j, "_clin.feat-", meta.name), obtain.bootstrap.metrics(metric.data, c("NPO", "Live_birth"))))
        }
      }
    }
  }
}

```

```{r, echo=F}

# blah.df <- readRDS("../results/kfold-panel-by-validation-split-internal/6/model-ridge_panel-val-6-lasso_max-iter-10_k-10_meta-_filter-0.05_class-hcg/combine-42.rds")
# blah.df[[1]]$preds$true
# obtain.bootstrap.metrics(blah.df)

```


# Features

```{r, echo=F}

gtf <- readRDS('../data/gtf.rds')
gtf.df <- data.frame(gene_id=gtf$gene_id, gene_name=gtf$gene_name, gene_type=gtf$gene_type) %>% distinct
gtf.df.protein.coding <- dplyr::filter(gtf.df, gene_type=='protein_coding')
gene.df <- gtf.df %>% tibble::column_to_rownames('gene_id') %>% dplyr::mutate(entrez=suppressWarnings(suppressMessages(mapIds(org.Hs.eg.db, keys = gsub('\\.[0-9]+', '', gtf.df$gene_id),
                                                                                                                              column = "ENTREZID", keytype = "ENSEMBL"))))
```

## 0.05 filter

```{r, echo=F}

filter.0.05 <- readRDS("../results/feature-selection/lasso-by-validation-split/6/max-iter-1000/prop-sels.rds")
gene.df$gene_name[match(filter.0.05, rownames(gene.df))]

```


## 0.1 filter

```{r, echo=F}

filter.0.1 <- readRDS("../results/feature-selection/lasso-by-validation-split-0.1-prop/6/max-iter-1000/prop-sels.rds")
gene.df$gene_name[match(filter.0.1, rownames(gene.df))]
```

# Summary

## Internal validation

```{r, echo=F, results="asis"}
rownames(internal.df) <- NULL
print(internal.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered"))
```

## External validation

```{r, echo=F, results="asis"}
rownames(sum.df) <- NULL
print(sum.df %>% kbl(format="html", align="c") %>% kable_styling(full_width = F, position="left", bootstrap_options="bordered"))
```

